---
title: "Analysis Report"
author: "MG"
date: "Modified: `r Sys.Date()`"
output: 
  bookdown::pdf_book: default
  bookdown::html_document2: default
bibliography: ["ref.bib"]
csl: ieee.csl
natbiboptions: round
link-citations: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# for not outputting echo
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
```

# Business Task Statement

The purpose of this study is to get some insights of space missions that
took place from 1957-2022. The data for this study is located at
[@maven_analytics] and contains 4,630 records regarding space missions
around the world for the period 10/4/1957 - 07/29/2022.

More specifically, this report finds answers to the following questions

- Launches: 
  - How many launches in total
  - How many successful/failed launches?
- Facilities: 
  - How many launches per facility?
  - Which countries launched space missions?
- Rockets:
  - How many rockets?
  - How many is active vs. retiered?
  - Which rockets were used the most?
  - Which rockets served the longest?

```{r loadlib}
# the default timeframe of the dataframe
LAUNCH_PERIOD="10/4/1957 - 7/29/2022"
LAUNCH_PERIOD_DOT=paste0(LAUNCH_PERIOD, ".")

# most intense color 
MOST_INTENSE_COLOR="#132B43"  # dark blue
LEAST_INTENSE_COLOR="#56B1F7" # light blue

# read the csv file
library(readr)
df <- read_csv('mission-all-fields.csv')
summary(df)

library(dplyr)
library(tidyverse)
```

````{=html}
<!-- If you want to make it visible despite was globally defined
```{r analysis, results='markup'}
summary(df)
```
-->
````

# Findings

Findings regard period `r LAUNCH_PERIOD`, unless stated otherwise.

<!-- ```{r, results='markup'} -->

```{r}
# we will be modifying df so making a copy
df_facility <- df
# concatenate Facility, State, Area and skip NA values
df_facility$Place <- apply(df[,3:5], 1, function(x) paste(na.omit(x), collapse = ", "))
# how many NA is per column in df_facility
sapply(df_facility, function(x) sum(is.na(x)))
#View(df_facility)

# how many facilities we have
print('Launch Facilities count:')
nrow(df_facility %>% distinct(Place))

# how many locations do we have
print('No. of countries with launch facilities:')
nrow(df_facility %>% distinct(Area))
```

```{r}
# for launches

# find total number of launches per place and their successes and failures
mission_status_df <- df_facility %>% group_by(Place) %>% 
  summarize(Total = n(), 
            Success = sum(MissionStatus == 1), 
            Failure = sum(MissionStatus == 2),
            PartialFailure = sum(MissionStatus == 3),
            PrelaunchFailure = sum(MissionStatus == 4),
            JustFailure = Total - Success) %>% 
  arrange(Place)

# add abbreviation for diagrams
mission_status_df$PlaceAbbrv  <- abbreviate(mission_status_df$Place, minlength = 45)

# have percentages ready for use
total_launches <- sum(mission_status_df$Total)

library(scales)

#' Compute the percentage
#' @param value Value to be presented as percentage
#' @param digi What precision, how many digits after decimal point
#' @return rounded value with digi digits after decimal point
percentage <- function(value, digi) {
  round(value*100, digits = digi)
}

# total percent success and total percent of failures
#mission_status_df$SuccessPercent <- percent(mission_status_df$Success/total_launches, accuracy = 0.1)
#mission_status_df$FailurePercent <- percent(mission_status_df$JustFailure/total_launches, accuracy = 0.1)
mission_status_df$SuccessPercent <- percentage(mission_status_df$Success/total_launches, 1)
mission_status_df$FailurePercent <- percentage(mission_status_df$JustFailure/total_launches, 1)
mission_status_df$TotalPercent <- percentage(mission_status_df$Total/total_launches, 1)
mission_status_df$RelSuccessPerc <- percentage(mission_status_df$Success/mission_status_df$Total, 1)
mission_status_df$RelFailurePerc <- percentage(mission_status_df$JustFailure/mission_status_df$Total, 1)

# find out total number of successes, failures and others
totals_df <- colSums(mission_status_df[,2:6])

```

- There were `r total_launches` launches in total. 
- There were `r nrow(df_facility %>% distinct(Place))` facilities
    capable of rocket launch.
- Facilities were located in `r nrow(df_facility %>% distinct(Area))`
  countries (or areas).

- `r percentage(41/nrow(df_facility %>% distinct(Place)),0)`% of facilities performed no more than 200 launches in total.

```{r}
hist(mission_status_df$Total, xlab = "Number of launches", 
     ylim=c(0,50),
     labels=TRUE, main = paste0(
  'Histogram of the total number of launches by facilities\nall over ',
  'the world for the period ', LAUNCH_PERIOD_DOT))
```
    
## Facilities

- A country can have several launching facilities.

```{r}
# get the number of facilities per country
countries_facilities_df <- df_facility %>% select(Area, Facility) %>% distinct(Facility, Area) %>% group_by(Area) %>% summarize(FacilityCount = n()) %>% arrange(desc(FacilityCount))

# add the percentage column
facilities_count <- sum(countries_facilities_df$FacilityCount)
countries_facilities_df$Percent <- percentage(countries_facilities_df$FacilityCount / facilities_count, 0)

hist(countries_facilities_df$FacilityCount, xlab = "Number of facilities per country", 
     labels=TRUE, 
     xlim = c(0, 20),
     ylim = c(0, 20),
     main = paste0('Histogram of the number of facilities per country',
                   '\nfor the period ', LAUNCH_PERIOD_DOT))
axis(side = 1, at = seq(0,20, 1))
```

- Facilities were located in `r nrow(countries_facilities_df)` countries.

- There were two locations with undefined country of ownership:
    Yellow Sea, and Barents Sea, where rockets were
    launched from submarines (Barents Sea) or barges (Yellow Sea).
    
```{r}

library(ggplot2)
ggplot(data=countries_facilities_df, aes(reorder(Area, Percent), Percent, fill=Percent)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = Percent), vjust=0.5, hjust=-0.3) +
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  xlab('Facility Location') +
  ylab('% of Total Number of Facilities') +
  coord_flip() +
  labs(fill='% Facilities\nLocated\nin\nCountry', 
       title=paste0('Percent of facilities located in a country for the period\n',
                    LAUNCH_PERIOD, '.'),
       subtitle=paste0('Number of facilities: ', facilities_count, 
                      '.'))
```
  

```{r, results='markup'}
knitr::kable(countries_facilities_df, caption=paste0('Facilities per location/country for the period ', LAUNCH_PERIOD_DOT))
```


    
- Eight countries were used by more than one company. Top 5 most
    popular launching locations:

    -   USA (used by 21 companies),
    -   Kazakhstan (used by 11 companies),
    -   Russia (used by 9 companies),
    -   China (used by 8 companies),
    -   Japan (used by 4 companies).

```{r}
# count distinct companies that were launching in a given area
companies <- df %>%
  distinct(Company, Area) %>% 
  group_by(Area) %>% 
  drop_na() %>% 
  summarize(count_companies = n()) %>% 
  arrange(count_companies, Area)

# to see the geographical locations used for mission launches
# View(companies)

# to see companies launching in a geographical location
#View(df_facility %>% distinct(Company, Area) %>% group_by(Company) %>% arrange(Company, Area))


g <- ggplot(data=companies, aes(reorder(Area, count_companies), count_companies, fill=count_companies))
  g + 
  geom_bar(stat="identity") +
  geom_text(aes(label = count_companies), vjust=0.5, hjust=-0.3) +
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  xlab('Country of a Facility(-ies)') +
  ylab('No. of Companies') +
  coord_flip() +
  labs(fill='Companies\nCount', 
       title=paste('The number of companies launching in a given country\n(',              LAUNCH_PERIOD, ').'),
       subtitle='Yellow Sea and Barents Sea - launches from barges and submarines.')
```

## Companies

<!-- do ```{r, results='markup'} to show the R output -->

```{r}
# How many companies do we have
df %>% distinct(Company) %>% nrow()

# ------------------------------------------------------
# Do companies launch only in one geographical location?
# ------------------------------------------------------
require(tidyr)   # for drop_na
locations <- df %>%
    distinct(Company, Area) %>% 
    group_by(Company) %>% 
    drop_na() %>% 
    filter(n() >= 2) %>% 
    arrange(Company, Area)
#View(locations)

# how many companies launched in a single location 
# the numbers should add up with previous numbers
df %>% distinct(Company, Area) %>% 
  group_by(Company) %>% 
  drop_na() %>% 
  filter(n() < 2) %>% 
  arrange(Company, Area) %>%
  nrow()
```

```{r}
# view the companies that launched space missions in more than one
# country
#View( df %>%
#    distinct(Company, Area) %>% 
#    group_by(Company) %>% 
#    drop_na() %>% 
#    filter(n() >= 2) %>% 
#    arrange(Company, Area) )

# how many companies launched in a single country 
# the numbers should add up with previous numbers to the total
# number of companies
df %>% distinct(Company, Area) %>% 
  group_by(Company) %>% 
  drop_na() %>% 
  filter(n() < 2) %>% 
  arrange(Company, Area) %>%
  nrow()
```

- There were `r df %>% distinct(Company) %>% nrow()` companies that
  launched space missions.
- `r df %>% distinct(Company, Area) %>% group_by(Company) %>% filter(n() < 2) %>% nrow()` companies launched in a single country/location.
- There were twelve companies that launched space missions in two
  different countries.

## Launches


- Top 5 countries/locations (USA, Russia, Kazahstan, China, France) account for over 92% of global launches. There were `r total_launches` launches in total.
```{r}
# launches per location/country

launches_per_loc_df <- df_facility %>%
  group_by(Area) %>%
  summarize(Total = n(), Percent = percentage(Total/ total_launches, 1)) %>%
  arrange(desc(Total))

launches_per_loc_df %>% select (Area, Percent) %>% 
  ggplot(aes(reorder(Area, Percent), Percent, fill=Percent)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_text(aes(label = Percent), vjust=0.5, hjust=-0.3) +
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  xlab('Location/Country') +
  ylab('% of Total Launches') +
  ylim(0,35) +
  coord_flip() +
  labs(fill='Percent', 
       title=paste('Locations/countries w.r.t. the number of launches globally',
       ' for\nthe period ', LAUNCH_PERIOD_DOT),
       subtitle=paste('There were', total_launches,   'launches in total.'))
```

```{r, results='markup'}
knitr::kable(launches_per_loc_df, caption=paste0('Locations/countries launching space missions ordered by percentage of the global number of launches for the period ', LAUNCH_PERIOD_DOT))
```

-   Top 10 launching facilities taking into account the number of
    launches.

```{r}
mission_status_df %>%
  select(PlaceAbbrv, TotalPercent) %>%
  arrange(desc(TotalPercent)) %>%
  head(10) %>%
  ggplot(aes(reorder(PlaceAbbrv, TotalPercent), TotalPercent, fill=TotalPercent)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_text(aes(label = TotalPercent), vjust=0.5, hjust=-0.3) +
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  xlab('Facility') +
  ylab('% of Total Launches') +
  ylim(0,30) +
  coord_flip() +
  labs(fill='Percent', 
       title=paste('Top 10 facilities w.r.t. the total number of launches',
                   '\nfor the period', LAUNCH_PERIOD_DOT),
       subtitle=paste('There were', total_launches,   'launches in total.'))
```

-   90% (`r totals_df[2][1]`) of `r totals_df[1][1]` launches in total
    were successful.

```{r}
# install.packages("lessR")
library(lessR)

counts_c <- totals_df[2:5]
# Color palette
colors <- hcl.colors(length(counts_c), "Blues")
pc(counts_c, 
   data = counts_c,
   fill=colors,
   values_size = 0.95,
   main=paste0("Space missions launch successes and failures\nduring ", LAUNCH_PERIOD, '. There were ', total_launches, ' launches in total.'))
```

```{r, results='markup'}
knitr::kable(head(totals_df), caption=paste0('The number of successful and failed launches per facility for the period ', LAUNCH_PERIOD_DOT))
```


```{r}
# I needed to reshape the data frame by melting status and counts
# to make a plot showing how many mission successes and failures.
library(reshape2)

molten.mission_status_df <- mission_status_df
molten.mission_status_df <- melt(molten.mission_status_df, id = c("Place", "PlaceAbbrv"))
#View(molten.mission_status_df)
```

<!-- this shows the raw numbers regarding successes and failed launches per facility
     since I present the percentages I skip this diagram
```{r}
# raw numbers - the number of success and failed launches
molten.mission_status_df %>% filter(variable %in% c('Success', 'JustFailure')) %>% 
 ggplot(aes(reorder(PlaceAbbrv, value), value, fill=variable)) +
  geom_bar(position=position_dodge(), stat="identity") +
  xlab('Facility') +
  ylab('Count') +
  coord_flip() +
  labs(fill='Mission Status', 
       title=paste('The number of success and failed launches\nper facility for the period', LAUNCH_PERIOD_DOT)) +
  scale_fill_discrete(labels = c("Success", "Failure"))
```
-->

- Percentage of successful and failed launches is shown below:

```{r}
# mission status percent
molten.ms_percent <- mission_status_df[,c("PlaceAbbrv","SuccessPercent",  "FailurePercent")]
names(molten.ms_percent) <- c("Place", "Success", "Failure")
#View(molten.ms_percent)

# now melt it
molten.ms_percent <- melt(molten.ms_percent[], id = c("Place"), value.name="Percent", variable.name = c("MissionStatus"))
#View(molten.ms_percent)

# install.packages("tidytext")
library(tidytext)
# now show it
molten.ms_percent %>% 
  ggplot(aes(reorder(Place, Percent), Percent, fill=MissionStatus)) +
  geom_bar(position=position_dodge(), stat="identity") +
  coord_flip() +
  xlab('Facility') +
  ylab('Percentage [%]') +
  labs(fill='Mission Status', 
       title=paste('Percentage of successful and failed launches\nper facility for the period', LAUNCH_PERIOD_DOT), 
       subtitle = paste('There were ', total_launches, ' launches in total.')) +
  scale_fill_discrete(labels = c("Success", "Failure"))
```

- Top 10 launching facilities with the highest number of space mission launches.

```{r, results='markup'}
knitr::kable( mission_status_df %>%
  select(Place, TotalPercent, Total, RelSuccessPerc, RelFailurePerc) %>%
  arrange(desc(Total)) %>%
  head(n=10),
  caption=paste0('Top 10 launching facilities for the period ', LAUNCH_PERIOD_DOT))
```

- Top 10 facilities with the highest number of launches 
  ordered by successful launches relative to the total
  number of launches a given facility performed. Kennedy Space Center had
  the highest relative success value, however, it performed only 4.4%
  of all global launches.
  
```{r}
# most launching facilities ordered by success in a descending order
most_launches_most_success <- mission_status_df %>%
  select(Place, TotalPercent, Total, RelSuccessPerc, RelFailurePerc) %>%
  arrange(desc(Total)) %>%
  head(n=10) %>%
  arrange(desc(RelSuccessPerc))

# position_dodge() put positions side by side
most_launches_most_success %>% 
  ggplot(aes(reorder(Place, RelSuccessPerc), RelSuccessPerc, fill=TotalPercent)) +
  geom_bar(position=position_dodge(), stat="identity") +
  coord_flip() +
  xlab('Facility') +
  ylab('Successful Launches [%]') +
  labs(fill='% Of All\nLaunches\nGlobally', 
       title=paste(
         'Top 10 most launching facilities',
         '\nordered by percentage of successful launches',
         '\nrelative to the total number of launches',
         '\na given facility performed',
         '\nfor the period', LAUNCH_PERIOD_DOT),
       subtitle = paste(
         'The color represents percentage of launches',
         '\na given facility performed w.r.t. all launches',
         '\nperformed globally, ', total_launches, 'in total.')) +
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  #scale_fill_viridis(option="E", na.value = "gray90", direction=-1) +
  scale_y_continuous(breaks = seq(from=0, to=100, by=10))
```
  
```{r, results='markup'}
knitr::kable( most_launches_most_success %>%
  select(Place, RelSuccessPerc, TotalPercent, Total) %>%
  arrange(desc(RelSuccessPerc)) %>%
  head(n=10),
  caption=paste0('Facilities with most launches globally ordered by successful launches relative to the number of launches a given facility performed for the period ', LAUNCH_PERIOD_DOT))
```

```{r}
# how many launches per year
#install.packages("lubridate")
library(lubridate)

# get the year from the date
# for some reason this does not work but it works in a R script
# #years_vec <- format(as.Date(df$Date, format="%Y-%m-%d"),"%Y")
years_vec <- year(mdy(df$Date))

# table() will get you the frequency table
years_df <- as_data_frame(table(years_vec))

# rename the columns names
colnames(years_df) <- c('Year', 'LaunchesCount')

# 'Year' is factor (see class(years_df$Year)), so it contains levels
# we want to convert it to numeric values and it is done via
# converting to a character and numeric
years_df$Year <- as.numeric(as.character(years_df$Year))

# add the percentage relative to the total launches
years_df$Percent <- percentage(years_df$LaunchesCount/total_launches,1)

# when is the maximum number of launches
max_number_of_launches <- years_df %>% filter( years_df$LaunchesCount == max(years_df$LaunchesCount))
# when the minimum number of launches occurred
min_number_of_launches <- years_df %>% filter( years_df$LaunchesCount == min(years_df$LaunchesCount))

# fast method for finding the second lowest; 
second_lowest <- sort(years_df$LaunchesCount,partial=2)[2]
second_min <- years_df %>% filter( years_df$LaunchesCount == second_lowest)

```
- Descriptive stats regarding launches w.r.t. launches per year for the period `r LAUNCH_PERIOD_DOT`:

```{r}
# because I do not want it to be visible so I first I have to prepare
# and transform the data, then use it in a table
launches_count_summary_df <- data.frame(unclass(summary(years_df$LaunchesCount)), check.names = FALSE)
colnames(launches_count_summary_df) <- c('LaunchesCount')
launches_count_summary_df$LaunchesCount <- round(launches_count_summary_df$LaunchesCount, 1)
launches_count_summary_df <- t(launches_count_summary_df)
```
```{r, results='markup'}
knitr::kable( launches_count_summary_df,
  caption=paste0('Stats for the launches for the period ', LAUNCH_PERIOD_DOT))
```

- The number of launches was the highest in 1966-1978 (around 100 per year), and 2016-2022 (around 100 per year). The lowest number of launches per year was performed in 2001-2015 that ranged from `r min(years_df[years_df$Year >= 2001 & years_df$Year <= 2015,]$Percent)`% of launches counted globally (`r min(years_df[years_df$Year >= 2001 & years_df$Year <= 2015,]$LaunchesCount)`) to `r max(years_df[years_df$Year >= 2001 & years_df$Year <= 2015,]$Percent)`% (`r max(years_df[years_df$Year >= 2001 & years_df$Year <= 2015,]$LaunchesCount)`). 
  
```{r}
# percentages w.r.t. the global number of launches
years_df %>% 
  ggplot(aes(x=Year, y=Percent, fill=Percent)) +
  geom_bar(position=position_dodge(), stat="identity") +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1, size=6)) +
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) + 
  ylab('% of Launches Globally') +
  labs(fill='% of\nLaunches\nGlobally', 
       title=paste(
         'Percentage of launches per year globally for the period', LAUNCH_PERIOD_DOT), 
       subtitle=paste('There were', total_launches, 'launches in total.'))


# for absolute numbers you can use this
# years_df %>% 
#  ggplot(aes(x=Year, y=LaunchesCount, fill=LaunchesCount)) +
#  geom_bar(position=position_dodge(), stat="identity") +
#  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1, size=6)) +
#  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) + 
#  ylab('No. of Launches') +
#  labs(fill='Number of\nLaunches', 
#       title=paste(
#         'The number of launches per year globally for the period', LAUNCH_PERIOD_DOT))
```

- The highest number of launches was in
  `r max_number_of_launches[1]` (`r max_number_of_launches[2]`) and it accounted for `r percentage(max_number_of_launches[2]/total_launches,1)`% of global launches. 
```{r}
most_launches_per_year <- years_df %>%
  arrange(desc(LaunchesCount), desc(Year)) %>%
  head(10)
# without conversion to a character it treats is as numerical values
# and displays accordingingly at the year axis (i.e., with gaps)
most_launches_per_year$Year <- as.character(most_launches_per_year$Year)

most_launches_per_year %>% 
  ggplot(aes(reorder(Year, Percent), y=Percent, fill=Percent)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_text(aes(label = Percent), vjust=0.5, hjust=-0.2) +
  coord_flip() +
  #scale_y_continuous(breaks = seq(from=0, to=160, by=10)) +
  #scale_x_continuous(breaks = seq(from=0, to=100, by=10))
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  xlab('Year') +
  ylab('% of Launches Globally') +
  labs(fill='% of\nLaunches\nGlobally', 
       title=paste(
         'Top 10 years with the most launches a year for the period\n', LAUNCH_PERIOD_DOT),
       subtitle=paste('There were', total_launches, 'launches in total.'))
```

- Year `r min_number_of_launches[1]` observed the lowest
  number of launches, i.e.,  `r min_number_of_launches[2]` accounting
  for `r min_number_of_launches[3]`% of total launches for the
  period. However, 1957 started in April. The second lowest yearly number of launches was in 
  `r second_min[1]`, i.e., `r second_min[2]` (`r second_min[3]`% of all global launches).

```{r}
least_launches_per_year <- years_df %>%
  arrange(LaunchesCount, Year) %>%
  head(10)
# without conversion to a character it treats is as numerical values
# and displays accordingingly at the year axis (i.e., with gaps)
least_launches_per_year$Year <- as.character(least_launches_per_year$Year)

least_launches_per_year %>% 
  ggplot(aes(reorder(Year, Percent, decreasing=TRUE), y=Percent, fill=Percent)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_text(aes(label = Percent), vjust=0.5, hjust=-0.2) +
  coord_flip() +
  #scale_y_continuous(breaks = seq(from=0, to=160, by=10)) +
  #scale_x_continuous(breaks = seq(from=0, to=100, by=10))
  scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  xlab('Year') +
  ylab('% of Launches Globally') +
  labs(fill='% of\nLaunches\nGlobally', 
       title=paste(
         'Bottom 10 years with the least launches a year for the period\n', LAUNCH_PERIOD_DOT),
       subtitle=paste('There were', total_launches, 'launches in total.'))
```

## Rockets

```{r}
# how many and what type of rockets are used
# how many times each rocket was used and what is its status (retired or not)
# how long it served

# first deal with the date and convert it to the right format
# for some reason the code that works fine in the R script
# does not work in a R Markdown chunk so here baby steps
rockets_df <- df %>%
  select(Rocket, RocketStatus, Date) 
# convert the Date column from character to the Date
rockets_df$Date <- mdy(rockets_df$Date)

# now deal with other stuff related to a date column
# the OperationInYears - is an approximate number (the year is assumed to
# have 365 days)
rockets_df <- rockets_df %>%
  group_by(Rocket, RocketStatus) %>% 
  summarize(LaunchCount = n(), FirstLaunch = min(Date), LastLaunch = max(Date),
            OperationInDays = length(seq(from = min(Date), to = max(Date), by = 'day')) - 1,
            OperationInYears = round(OperationInDays / 365, digits = 1)) %>% 
  arrange(desc(LaunchCount), desc(Rocket))

rockets_df$CountPerc <- percentage(rockets_df$LaunchCount / sum(rockets_df$LaunchCount), 1)

# hack to make nice labels in PieChart - I need to add a label
# otherwise I have 0 and 1, so here we go
# add the column with labels
rockets_df$RocketStatusName <- with(rockets_df, ifelse(RocketStatus == 0, "Retired", "Active"))
```

- There were total `r nrow(rockets_df)` rockets used for launching. Most rockets
  (`r percentage(359/nrow(rockets_df), 0)`%) performed no more than 50 launches.
```{r}
hist(rockets_df$LaunchCount, xlab = "No. of Launches", xlim=c(0,500), ylim=c(0,400), labels=TRUE, main = paste0(
  'Histogram of number of launches per rocket for the period\n', LAUNCH_PERIOD_DOT,
  ' Total number of rockets: ', nrow(rockets_df), '\nTotal number of launches: ',
  nrow(df), '.'))
```

```{r}
# because I do not want it to be visible so I first I have to prepare
# and transform the data, then use it in a table
rockets_summary_df <- data.frame(unclass(summary(
  rockets_df[c('LaunchCount', 'FirstLaunch', 'LastLaunch', 'OperationInYears')])), check.names = FALSE)

#colnames(rockets_summary_df)[1] <- 'LaunchCount'
rownames(rockets_summary_df) <- c('Min', '1st Qu', 'Median', 'Mean', '3rd Qu', 'Max')
```
```{r, results='markup'}
knitr::kable( rockets_summary_df[c('Min', 'Median','Mean', 'Max'),],
  caption=paste0('Stats for rockets for the period ', LAUNCH_PERIOD_DOT, 'Total number of rockets used is ', nrow(rockets_df)) )
```

- The status of rockets: active vs. retired.

```{r}
# how many rockets are active vs retired
PieChart(RocketStatusName, data=rockets_df,
  fill=c("red","darkblue"),
   values_size = 0.95,
   values = "%",
   labels_cex = 1.5,
   cex = 1.5,
   main=paste("Active vs. retired rockets as on 7/29/2022",
   "\nfor the period", LAUNCH_PERIOD_DOT),
   add=paste("Total number\nof rockets = ", nrow(rockets_df)), x1=0, y1=0)
```


```{r}
# top 10 most popular rockets
top10_rockets_df <- rockets_df %>% 
  select(Rocket, LaunchCount, CountPerc, RocketStatusName, FirstLaunch, LastLaunch, OperationInYears) %>% 
  arrange(desc(LaunchCount)) %>% 
  head(10) 
```

- Rocket `r top10_rockets_df$Rocket[1]` was the most popular one with 
`r top10_rockets_df$LaunchCount[1]` launches which accounted for `r top10_rockets_df$CountPerc[1]` of total launches.
 
```{r}
top10_rockets_df %>% 
  ggplot(aes(reorder(Rocket, CountPerc), y=CountPerc)) +
  geom_bar(position=position_dodge(), stat="identity", aes(fill=factor(RocketStatusName), group=1), show.legend = TRUE) +
  theme(axis.text.x = element_text(angle=0, vjust=.5, hjust=1)) +
  scale_fill_discrete(labels = c("Active", "Retired")) +
  #scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  ylab('% of total no. of launches by a given rocket') +
  xlab('Rocket') + 
  labs(fill="Rocket Status", 
       title=paste('The most popular rockets of all times\nfor the period', LAUNCH_PERIOD_DOT),
       subtitle=paste0('Total number of rockets was ', nrow(rockets_df), '.')) +
  coord_flip()
```

```{r, results='markup'}
knitr::kable( top10_rockets_df,
  caption=paste0('Most used rockets of all times for the period ', LAUNCH_PERIOD_DOT, 'Total number of rockets was ', nrow(rockets_df), '.'))
```

```{r}
# top 10 rockets serving the longest
top10_longest_rockets_df <- rockets_df %>% 
  select(Rocket, LaunchCount, CountPerc, RocketStatusName, FirstLaunch, LastLaunch, OperationInYears) %>% 
  arrange(desc(OperationInYears)) %>% 
  head(10) 
```


- There were total `r nrow(rockets_df)` rockets used for launching. Most rockets
  (`r percentage(237/nrow(rockets_df), 0)`%) operated for no more than 5 years.
```{r}
hist(rockets_df$OperationInYears, xlab = "Operation in Years", 
     xlim=c(0,50), ylim=c(0,250), labels=TRUE, breaks=9,main = paste0(
  'Histogram of operational rocket lifespan for the period\n', LAUNCH_PERIOD_DOT,
  ' Total number of rockets: ', nrow(rockets_df), '.'))
```

- The longest serving rocket was `r top10_longest_rockets_df[1,][1]` which served
from `r top10_longest_rockets_df[1,][5]` and retired on `r top10_longest_rockets_df[1,][6]`, accounting for `r top10_longest_rockets_df[1,][7]` years of operation. The longest serving rocket still active as of 
the end of period for this dataset was `r top10_longest_rockets_df[3,][1]` with
`r top10_longest_rockets_df[3,][7]` years of operation.

```{r}
top10_longest_rockets_df %>% 
  ggplot(aes(reorder(Rocket, OperationInYears), y=OperationInYears)) +
  geom_bar(position=position_dodge(), stat="identity", aes(fill=factor(RocketStatusName), group=1), show.legend = TRUE) +
  theme(axis.text.x = element_text(angle=0, vjust=.5, hjust=1)) +
  scale_fill_discrete(labels = c("Active", "Retired")) +
  #scale_fill_continuous(high = MOST_INTENSE_COLOR, low = LEAST_INTENSE_COLOR) +
  ylab('Years in Operation') +
  xlab('Rocket') + 
  labs(fill="Rocket Status", 
       title=paste('Rockets which served the longest for the period\n', LAUNCH_PERIOD_DOT),
       subtitle=paste0('Total number of rockets was ', nrow(rockets_df), '.')) +
  coord_flip()
```

```{r, results='markup'}
knitr::kable( top10_longest_rockets_df,
  caption=paste0('Rockets that served the longest for the period ', LAUNCH_PERIOD_DOT, 'Total number of rockets was ', nrow(rockets_df), '.'))
```



# Preparing Data

The original data is located at [@maven_analytics]. The data is
organized in two .csv files:

-   `space_missions.csv` - the actual data
-   `space_missions_data_dictionary.csv` - description of fields

| Data Structure | No. of Fields | No. of Records | Date Added |
|:--------------:|:-------------:|:--------------:|:----------:|
|  Single table  |       9       | 4,630 + header | 08/03/2022 |

## Data Structure

Each observation consists of 9 attributes that describe: Company,
Location, Date, Time, Rocket, Mission, RocketStatus, Price,
MissionStatus.

-   Company - a text field that can includes spaces, variable length
-   Location - a text field, variable length, includes: the site, place,
    state, country. State and country is for USA. Tokens separated by
    comma.
-   Date - m/d/yyyy; m and d can be a single or double digit
-   Time - h:m:ss; h and m can be a single or double digit; according
    description UTC
-   Rocket - a text that can have multiple tokens
-   Mission - a text that can contain multiple tokens; letter encoding
    issues
-   RocketStatus - textual one word (Retired, Active)
-   Price - numeric, often missing
-   MissionStatus - (Success, Failure, Partial Failure, Prelaunch
    Failure)

## Data Credibility

The data is provided by a respected website devoted to Data Analytics
[@maven_analytics]. I checked and confirmed from other sources a couple
randomly selected observations:

-   Falcon 9 Block 5 \| Nilesat-301 \| SpaceX \| June 8th, 2022
-   Martin Marietta SLC-4E, Vandenberg AFB, California, USA 11/8/1991
    7:07:00 Titan IV(403)A SLDCOM & NOSS 3x

## Data Bias

To rule out the bias in the data, the research should be performed to
verify if all space missions were indeed included within the specified
timeline, i.e., 1957-2022. My assumptions is that the data is collected
and presented in a fair manner.

## License

I have not found any license on how to use the data. I have reached out
to the staff and they said that "they [datasets] are free to use as you
wish".

## Potential Issues

-   Some data is missing. Especially, the price field.
-   Mission field has different encoding for letters.
-   Location field probably needs to be separated.

## Preparation Methodology

I used Microsoft Excel to visually inspect the data and address the
issues.

# Data Processing

I could not use pivot tables for the dataset due to a MS Excel memory
issue: when trying to create a pivot table from the data csv file with
4,630 observations and 9 attributes, the message showed up:

"Excel cannot complete this task with available resource. Choose less
data or close other applications."

## Data Cleaning

Data cleaning has been performed in MS Excel and R. Mostly manually with
the help of MS Excel tools such as Find/Replace, Missing Values Filter,
and R to find out encoding issues (writing functions that report
issues).

It resulted in creation of a file named `space_mission.xlsx` that I used
for further data analysis.

### Field: Location

-   Algeria: Corrected the country for launch for - Brigitte, Hammaguir,
    Algeria, France. Removed France because the location is in Algeria,
    not in France (4 occurrences).
-   Kiribati: Kiritimati Launch Area has Pacific Ocean, although it
    belongs to the Republic of Kiribati (36 occurrences)
-   Barents Sea: it was launched in the Barents Sea Launch Area by
    Russian submarines (3 occurrences).
-   Yellow Sea

### Field: Date

-   Added missing Date for "Rocket Lab Rocket Lab LC-1A, Mahia
    Peninsula, New Zealand 0:02:00 Electron/Curie\
    A Data With Destiny"

### Field: Time

-   There are 127 missing values. I have tried to find first ten of them
    to no avail.

### Field: MissionStatus

I added a numerical field that I used in csv later on:

| Original MissionStatus | Numerical MissionStatus |
|:----------------------:|:-----------------------:|
|        Success         |            1            |
|        Failure         |            2            |
|    Partial Failure     |            3            |
|   Prelaunch Failure    |            4            |

### Checks

Total number of observations: 4630

| Field                  | Missing | Range/Values                                         |
|:------------------------|-----------------:|:----------------------------|
| Company                |       0 | len: 2-16                                            |
| Location               |      \- | \-                                                   |
| Site                   |      23 | len: 0-23                                            |
| Facility               |       3 | len: 0-49                                            |
| State                  |    2839 | len: 0-16                                            |
| Area                   |       0 | len: 3-13                                            |
| Date                   |       0 | 10/4/1957 - 7/29/2022                                |
| Time                   |     127 | 0:00-23:59                                           |
| Rocket                 |       0 | len: 3-24                                            |
| Mission                |       0 | len: 3-51                                            |
| RocketStatus           |       0 | Retired, Active                                      |
| Price (millions of \$) |    3365 | 2.5 - 5,000                                          |
| MissionStatus          |       0 | Success, Failure, Partial Failure, Prelaunch Failure |

### Result

I cleaned manually names from all characters that might be problematic
for the databases or further processing.

I created the `space-mission.xlsx` file and exported `missions.csv` and
`places.csv` that can be used in further processing.

-   `places.csv` - contains part of the original location, i.e.,
    facility, state, area
-   `missions.csv` - contains RocketStatus and MissionStatus changed to
    numbers, it also has an additional field called ID of the mission.
-   `missions-all-fields.csv` - contains numeric values of RocketStatus
    and MissionStatus, the location has four elements

# References
